---
title: "RP2"
author: "GROUP 3"
date: "2025-01-22"
output: html_document
---

```

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.height = 8)

```

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(printr)
library(readr)
library(car)
library(profvis)

require(readxl)
require(jpeg)
require(ggimg)

```
### Input and preparation

###### Reading participant data from the RP2 folder.

```{r}

# Get list of all files in RP2 directory
rp2_files <- list.files(".", pattern = "*.csv", full.names = TRUE) #enter working directory

```

```{r}

# Define function to load and clean files from RP2
read_rp2_files <- function(file_path) {
  read_csv(file_path, 
           col_types = cols(
             recording_start_time = col_double(),
             current_phase = col_character(),
             value = col_double()
           )) %>%
    filter(!is.na(value)) # Remove rows without valid participant data
}
```

```{r load-data}
# Load and combine participant data
participant_data <- bind_rows(lapply(rp2_files, read_rp2_files), .id = "file_id")

# Inspect combined data
head(participant_data)

```

### Comparing Participant Data

###### Aggregating and comparing data across participants.

```{r}
# Summarize participant data by file and phase
summary_data <- participant_data %>%
  group_by(file_id, current_phase) %>%
  summarize(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE),
    min_value = min(value, na.rm = TRUE),
    max_value = max(value, na.rm = TRUE),
    .groups = "drop"
  )

# Inspect summary statistics
print(summary_data)
```

### Visualizing Participant Comparisons

###### Plotting participant data for comparison.

```{r, fig.height = 8, fig.width = 10}
# Boxplot of values by phase and file
participant_data %>%
  ggplot(aes(x = current_phase, y = value, fill = file_id)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Participant Data Comparison",
    x = "Phase",
    y = "Value",
    fill = "Participant"
  )
```


```{r, fig.height = 8, fig.width = 10}
# Bar graph of mean values by phase for each file
summary_data %>%
  ggplot(aes(x = current_phase, y = mean_value, fill = file_id)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(
    title = "Mean Value by Phase",
    x = "Phase",
    y = "Mean Value",
    fill = "Participant"
  )
```
###### Exploring the variability between participants in each phase
```{r}
participant_variability <- participant_data %>%
  group_by(file_id, current_phase) %>%
  summarize(
    median_value = median(value, na.rm = TRUE),
    interquartile_range = IQR(value, na.rm = TRUE),
    #missing_count = sum(is.na(value)),
    .groups = "drop"
  )
print(participant_variability)
```
```{r}
participant_variability %>%
  pivot_longer(cols = c(median_value, interquartile_range), 
               names_to = "metric", 
               values_to = "value") %>%
  ggplot(aes(x = current_phase, y = value, fill = file_id)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ metric, scales = "free") +
  theme_minimal() +
  labs(
    title = "Participant Variability Metrics by Phase",
    x = "Phase",
    y = "Metric Value",
    fill = "Participant"
  )
```

```{r, fig.height = 8, fig.width = 10}
participant_data %>%
  ggplot(aes(x = current_phase, y = value, fill = current_phase)) +
  geom_boxplot() +
  facet_wrap(~ file_id, scales = "free") +
  theme_minimal() +
  labs(
    title = "Phase-wise Boxplot for Each Participant",
    x = "Phase",
    y = "Value",
    fill = "Phase"
  )
```

#### Distributions of values for participants in each phase.
```{r, fig.height = 8, fig.width = 10}
participant_data %>%
  ggplot(aes(x = value, fill = current_phase)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ file_id) +
  theme_minimal() +
  labs(
    title = "Value Distribution by Participant and Phase",
    x = "Value",
    fill = "Phase"
  )
```


###### Performing paired t-tests or repeated-measures ANOVA to identify significant differences between phases.

```{r}
# Paired t-test with Debugging
intermediate_table <- participant_data %>%
  filter(current_phase %in% c("Phase 1- Baseline", "Phase 2")) %>%
  group_by(file_id, current_phase) %>%
  summarize(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  spread(key = current_phase, value = mean_value)

# Perform paired t-test
paired_test <- t.test(intermediate_table$`Phase 1- Baseline`, intermediate_table$`Phase 2`, paired = TRUE)
print(paired_test)
```

```{r}
install.packages("plotly")

library(plotly)
```

```{r}
ggplotly(
  participant_data %>%
    ggplot(aes(x = current_phase, y = value, color = file_id)) +
    geom_boxplot() +
    theme_minimal() +
    labs(
      title = "Interactive Participant Comparison",
      x = "Phase",
      y = "Value",
      color = "Participant"
    )
)
```

```{r}
correlation_data <- participant_data %>%
  group_by(file_id) %>%
  summarize(mean_baseline = mean(value[current_phase == "Phase 1- Baseline"], na.rm = TRUE),
            mean_phase2 = mean(value[current_phase == "Phase 2"], na.rm = TRUE)) %>%
  na.omit()

cor_test <- cor.test(correlation_data$mean_baseline, correlation_data$mean_phase2)
print(cor_test)
```


```{r}
# Paired t-test with Debugging
intermediate_table <- participant_data %>%
  filter(current_phase %in% c("Phase 2", "Phase 3")) %>%
  group_by(file_id, current_phase) %>%
  summarize(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  spread(key = current_phase, value = mean_value)

# Perform paired t-test
paired_test <- t.test(intermediate_table$`Phase 2`, intermediate_table$`Phase 3`, paired = TRUE)
print(paired_test)
```

```{r}
# One-way repeated-measures ANOVA (comparing all phases)

anova_data <- participant_data %>%
  group_by(file_id) %>%
  mutate(id = as.factor(row_number())) %>%
  ungroup()

#anova_model <- aov(value ~ current_phase + Error(id/current_phase), data = anova_data)

#summary(anova_model)
```

### Saving Cleaned Data
###### Exporting the cleaned and combined dataset for further analysis.

```{r}

save(participant_data,summary_data, intermediate_table, anova_data, file = "RP2.Rda")

```
