---
title: "RP2"
author: "GROUP 3"
date: "2025-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.height = 8)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(printr)
library(readr)
library(fmsb)

require(readxl)
```

### Input and preparation
###### Get list of all files in RP2 directory
```{r}
rp2_files <- list.files("C:/Users/mycom/Downloads/RP2/SENSOR_DATA", pattern = "*.csv", full.names = TRUE)
```

###### Create a reusable function to load and preprocess individual CSV files.
```{r}
# Define function to load and clean files from RP2
read_rp2_files <- function(file_path) {
  read_csv(file_path, 
           col_types = cols(
             recording_start_time = col_character(),
             current_phase = col_character(),
             value = col_double()
           )) %>%
    rename_all(~str_trim(.)) %>% # Standardize column names to prevent errors
    filter(!is.na(value))       # Remove rows without valid participant data
}
```

###### Load all participant data files, process them using the `read_rp2_files` function, and combine the resulting data frames into a single dataset.
```{r load-data}
participant_data <- bind_rows(lapply(rp2_files, read_rp2_files), .id = "file_id")

# Rename 'value' column to 'score' to reflect consistency with NASA TLX
participant_data <- participant_data %>% rename(score = value)

# Inspect combined data
summary(participant_data)
head(participant_data)
```

### Comparing Participant Data
###### Generate summary statistics for each participant and phase combination. Summarize key statistics for the "score" variable within each combination of participant and phase.
```{r}
summary_data <- participant_data %>%
  group_by(file_id, current_phase) %>%
  summarize(
    mean_value = mean(score, na.rm = TRUE),
    sd_value = sd(score, na.rm = TRUE),
    min_value = min(score, na.rm = TRUE),
    max_value = max(score, na.rm = TRUE),
    .groups = "drop"
  )

# Inspect summary statistics to confirm that statistics have been calculated correctly.
head(summary_data)
```

### Visualizing Participant Comparisons
###### To create visualizations that facilitate comparisons of participant data across phases. The visualizations help to identify patterns, outliers, and trends within the data.

```{r, fig.height = 8, fig.width = 10}
# Boxplot of values by phase and participants
participant_data %>%
  ggplot(aes(x = current_phase, y = score, fill = file_id)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Participant Data Comparison",
    x = "Phase",
    y = " Sensor Readings",
    fill = "Participant"
  )
```

###### Bar graph to visualize the mean for each phase across all participants
```{r, fig.height = 8, fig.width = 10}
summary_data %>%
  ggplot(aes(x = current_phase, y = mean_value, group = file_id, color = file_id)) +
  geom_line(size = 0.7) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(
    title = "Trends in Sensor Readings Across Phases",
    x = "Phase",
    y = "Avg Sensor Value",
    color = "Participant"
  )
```

##### Participant Level Analysis: Variability
###### Exploring the variability between participants in each phase
```{r, fig.height = 8, fig.width = 10}
participant_variability <- participant_data %>%
  group_by(file_id, current_phase) %>%
  summarize(
    median_value = median(score, na.rm = TRUE),
    interquartile_range = IQR(score, na.rm = TRUE),
    .groups = "drop"
  )

participant_data %>%
  ggplot(aes(y = score, fill = current_phase)) +
  geom_boxplot() +
  facet_wrap(~ file_id, strip.position = "bottom") +
  theme_minimal() +
  theme(strip.placement = "outside", strip.text.x = element_text(size = 10), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(
    title = "Within-Participant Trends: Sensor Readings in each phase",
    x = "Participant",
    y = "Sensor Value",
    fill = "Phase"
  )
```

#### Distributions of values for participants in each phase.
```{r, fig.height = 8, fig.width = 10}
participant_data %>%
  ggplot(aes(x = score, fill = current_phase)) +
  geom_density(alpha = 0.9) +
  facet_wrap(~ file_id) +
  theme_minimal() +
  labs(
    title = "Value Distribution by Participant and Phase",
    x = "Sensor Value",
    fill = "Phase"
  )
```

##### Load NASA TLX data
```{r}
nasa_tlx_scores <- read_csv("C:/Users/mycom/Downloads/RP2/NASATLX/NASA_TLX_Scores.csv") %>%
  mutate(
    current_phase = factor(current_phase, levels = c("Phase 2", "Phase 3")),
    file_id = as.character(file_id)  # Align with participant_data
  )
```

```{r}
# Function to prepare data for radar charts
prepare_radar_data <- function(data, participant_id, phase) {
  data %>%
    filter(file_id == participant_id & current_phase == phase) %>%
    select(mental_demand, physical_demand, temporal_demand, 
           performance, effort, frustration) %>%
    {rbind(rep(100, ncol(.)), rep(0, ncol(.)), .)} # Add max and min rows for scaling
}

# Generate radar charts for each participant and phase
unique_participants <- unique(nasa_tlx_scores$file_id)
unique_phases <- unique(nasa_tlx_scores$current_phase)

for (participant_id in unique_participants) {
  for (phase in unique_phases) {
    radar_data <- prepare_radar_data(nasa_tlx_scores, participant_id, phase)
    if (nrow(radar_data) > 2) {  # Ensure data is present
      radarchart(
        radar_data,
        axistype = 1,
        pcol = "blue", pfcol = rgb(0.2, 0.5, 0.5, 0.5), plwd = 2,
        cglcol = "grey", cglty = 1, axislabcol = "grey",
        caxislabels = seq(0, 100, 25), vlcex = 0.8
      )
      title(main = paste("Participant:", participant_id, "- Phase:", phase))
    }
  }
}
```


```{r}
# Merge sensor and NASA TLX data
combined_data <- participant_data %>%
  filter(current_phase %in% c("Phase 2", "Phase 3")) %>% # Filter for Phase 2 and 3 only
  group_by(file_id, current_phase) %>%
  summarize(
    mean_value = mean(score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(nasa_tlx_scores, by = c("file_id", "current_phase"))
```

#### Standardize both metrics using Z-Score
```{r}
standardized_data <- combined_data %>%
  mutate(
    standardized_mean_sensor_value = scale(mean_value)[, 1],  # Standardize sensor readings
    standardized_NASATLX_score = scale(score)[, 1]            # Standardize NASA TLX scores
  )

# Step 2: Reshape the data for combined plotting
long_standardized_data <- standardized_data %>%
  pivot_longer(
    cols = c(standardized_mean_sensor_value, standardized_NASATLX_score),
    names_to = "metric",
    values_to = "standardized_value"
  )
```

```{r, fig.height = 8, fig.width = 10}

participants <- unique(long_standardized_data$file_id)

# Create a single plot comparing participants on the x-axis
long_standardized_data %>%
  ggplot(aes(x = file_id, y = standardized_value, fill = metric)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +  # Add black outlines to bars
  facet_wrap(~ current_phase) +
  theme_minimal() +
  labs(
    title = "Comparison of Standardized NASA TLX and Sensor Readings Across Participants",
    x = "Participant",
    y = "Standardized Value (Z-Score)",
    fill = "Metric"
  )

```
```

### Saving Cleaned Data
###### Exporting the cleaned and combined dataset for further analysis.
```{r}
save(participant_data, summary_data, file = "RP2.Rda")
